---
title: "Chembela ANOVA lab"
author: "Hardlife Chinwa"
date: "2025-03-04"
output: html_document
---
###set the working directory

```{r}
setwd("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/Chembela")
```

###load the data, can use the import function

```{r}
radial=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/Chembela/design_lab.csv")
```
###attach the data
```{r}
attach(radial)
```

```{r}
library(ggplot2)
library(agricolae)
library(dplyr)
library(patchwork)
```

```{r}
radial$Replication=as.factor(radial$Replication)
radial$trt=as.factor(radial$trt)
radial$Day=as.factor(radial$Day)
```

###check for the normality of data
```{r}
shapiro.test(radial$Radial_growth)
```

###plot a normality line
```{r}
library(tidyselect)
library(ggplot2)
qqnorm(radial$Radial_growth)
qqline(radial$Radial_growth)
qqplot(radial$trt,radial$Radial_growth)
```

# Using to explore the distribution of a variable
```{r}
ggplot(radial, aes(sample = radial$Radial_growth)) +
  stat_qq() +
  stat_qq_line()
ggplot(radial, aes(sample = Radial_growth, colour = factor(trt))) +
  stat_qq() +
  stat_qq_line()+
  theme_bw()+
  theme(element_blank())
```

###compute the model lmer
```{r}
library(lme4)
library(lmerTest)
library(agricolae)
```

```{r}
rad=lm(Radial_growth~trt*Day+Replication,radial)
```

```{r}
rads=anova(rad)
```

```{r}
library(ggplot2)
library(agricolae)
library(dplyr)
library(patchwork)
```

####Pod incubation period####
```{r}
value_max=radial %>% group_by(trt) %>% summarize(max_value=max(Radial_growth))
hsd=HSD.test(aov(Radial_growth~trt,data=radial),trt="trt",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
Severityplot=ggplot(data=radial,aes(reorder(x=trt,-Radial_growth),y=Radial_growth))+
  geom_boxplot(aes(fill=trt))+
                 geom_text(data=value_max,aes(x=trt,y=0.+
                                                            max_value,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("Radial growth (mm)")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
  theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.x = element_text(size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(legend.position = "none")
```

```{r}
Severityplot
```

###compute 2 way interaction plots
```{r}
interaction.plot(x.factor = radial$trt,
                 trace.factor =radial$Day,
                 response = Radial_growth,
                 fun = mean,
                 xlab = "trt",
                 ylab = "Radial_growth",
                 trace.label = "Day",
                 col = factor(trt),
                 lty = 4,
                 lwd=2.5,xtick = TRUE,
                 main="Radial_growth")
with(radial,{interaction.plot(trt,Day,Radial_growth)
  interaction.plot(Day,trt,Radial_growth,cex.axis=0.8)
  Stage=factor(Day,levels = sort.list(tapply(Radial_growth,Day,mean)))
  interaction.plot(Day,trt,Radial_growth,col = factor(trt),lty = 4)
  
})
```

# Type 3 anova table:
# Display tests/hypotheses for type 1, 2, and 3 ANOVA tables:
# (and illustrate effects of 'fractions' and 'names' arguments)
```{r}
(an <- anova(chin, type="3"))
T1=show_tests(anova(chin, type="1"))
T2=show_tests(anova(chin, type="2"), fractions=TRUE, names=TRUE)
T3=show_tests(an, fractions=TRUE)
```

